---
title: "Gender dynamics experiments"
params:
  experiment: 1
  subset_name: "Sop_features_non_cat"
author: "Eduardo García-Portugués, Carlos Vaquero"
date: '`r Sys.Date()`'
format:
  pdf:
    latex_engine: xelatex
    number_sections: yes
    keep-tex: false
fontsize: 10pt
urlcolor: blue
editor_options:
  chunk_output_type: console
---

```{=html}
<!--

# To reproduce experiments, run ./experiments.sh or execute the following block in R
exps <- 1:3
subsets <- c(
  "Sop_features_non_cat"
)
for (exp in exps) {
  for (s in subsets) {
    message("Rendering for experiment: ", exp, " and subset: ", s)
    quarto::quarto_render(
      input = "Models.qmd",
      execute_params = list(experiment = exp, subset_name = s),
      output_file = paste0("Experiment_", exp,"_", s, ".pdf")
    )
  }
}

--->
```

```{r setup-params}
#| echo: false
# Report parameters
subset_name <- params$subset_name
experiment <- params$experiment
message("Running experiment ", experiment, " with subset ", subset_name)
```

# Load libraries

```{r, libraries}
#| message: false
# Load libraries
library(knitr)
library(dplyr)
library(recipes)
library(rsample)
library(yardstick)
library(glmnet)
library(ggplot2)
library(caret)
library(tibble)
sessionInfo()
```

```{r, functions}
#| include: false
#| cache: false
knitr::opts_chunk$set(
  echo = TRUE, cache = FALSE, warning = FALSE, message = FALSE,
  fig.path = paste0("Models_files/Experiment_", params$experiment, "_",
                    params$subset_name, "/")
)

# Compute AUC
auc_binary <- function(y, p_hat) {

  roc_auc_vec(
    truth = y,
    estimate = drop(p_hat),
    event_level = "second"
  )

}

# CIs for logistic ridge estimation
logistic_ridge_ci <- function(X, y, beta_hat, lambda, level = 0.95) {

  # X: design matrix (n x p)
  # y: binary response (0/1)
  # beta_hat: vector of coefficients (length p + 1)
  # lambda: penalty parameter in "theory" scale: (X'X + 2 * lambda * I)
  #         If the input of lambda is that from glmnet, then set
  #         lambda = n * lambda_glmnet / 2
  # level: confidence level, default 0.95

  # Variables
  n <- nrow(X)
  p <- ncol(X)

  # V_hat matrix
  eta_hat <- drop(beta_hat[1] + X %*% beta_hat[-1])
  p_hat <- 1 / (1 + exp(-eta_hat))
  V_hat <- diag(p_hat * (1 - p_hat), nrow = n, ncol = n)

  # Asymptotic covariance from Le Cessie & Van Houwelingen (1992, p. 194)
  # https://www.jstor.org/stable/2347628?seq=4
  # Recall the 2 * lambda in the ridge penalty
  XtWX <- t(X) %*% V_hat %*% X
  XtWX_lambda_inv <- solve(XtWX + diag(2 * lambda, nrow = p, ncol = p))
  asymp_var <- XtWX_lambda_inv %*% XtWX %*% XtWX_lambda_inv

  # Standard errors and z-based CIs
  se <- sqrt(diag(asymp_var))
  zcrit <- qnorm(1 - (1 - level) / 2)
  ci_low <- beta_hat[-1] - zcrit * se
  ci_up <- beta_hat[-1] + zcrit * se

  # Return a tibble
  tibble(
    term = colnames(X) %||% paste0("x", seq_len(p)),
    estimate = beta_hat[-1],
    std_error = se,
    ci_low = ci_low,
    ci_up = ci_up,
    z_value = estimate / std_error,
    p_value = 2 * pnorm(-abs(z_value))
  )

}

# Extract features according to the types of subsets defined
get_features_subsets_voice <- function(df_total) {

  # 1. Drop unwanted columns by regex
  df_total <- df_total |>
    select(-matches("_win_|js_.*|m21.*|SoundSop|Count|Sum|Notes|SoundingMeasures|SopI|SopII|SopIII|PartSop_Syllables"),
           -matches("^PartSop_HighestNote($|_[A-Za-z0-9]+)"),
           -matches("^PartSop_LowestNote($|_[A-Za-z0-9]+)"))

  # 2. Keep only Sop* features (no target)
  df_total <- df_total |>
    select(matches("^Sop|^PartSop")) |>
    # remove only columns like PartSop_HighestNote_G5 and PartSop_LowestNote_C4
    select(-matches("^PartSop_HighestNote_.+|^PartSop_LowestNote_.+"))

  # 3. Feature subset definitions
  features_subsets <- list(
    Melody = df_total |>
      select(matches("Amb|Density|Trimmed|Semitones|AbsoluteIntervallic|AscendingIntervals|AscendingIntervallic|StepwiseMotionDesc|PartSop_LowestNoteIndex|PartSop_HighestNoteIndex")) |>
      select(-matches("PartSop_RhythmInt")),

    Melody_extended = df_total |>
      select(matches("PartSop_Degree|IntervalA|IntervalM|IntervalP|Intervald|Intervalm|Amb|Density|Trimmed|Semitones|AbsoluteIntervallic|AscendingIntervals|AscendingIntervallic|StepwiseMotionDesc|PartSop_LowestNoteIndex|PartSop_HighestNoteIndex")) |>
      select(-matches("PartSop_RhythmInt")),

    Rhythm = df_total |>
      select(matches("DottedRhythm|AverageDuration|DoubleDottedRhythm|RhythmInt")),

    Lyrics = df_total |>
      select(matches("VoicePresence|SyllabicRatio")),

    All_sop_features = df_total |>
      select(matches("PartSop"))
  )

  # 4. Non-categorical version
  cols_to_drop <- features_subsets$All_sop_features |>
    select(matches("PartSop_Degree|PartSop_IntervalM|PartSop_IntervalP|PartSop_Intervald|PartSop_Intervalm|PartSop_LargestIntervalDesc_|PartSop_Interval*_Per|PartSop_MeanInterval|PartSop_SmallestIntervalAsc|PartSop_LargestIntervalDesc|PartSop_IntervalA.2_Per|PartSop_IntervalA.4_Per|PartSop_IntervalA.5_Per|PartSop_IntervalA1_Per|PartSop_IntervalA11_Per|PartSop_IntervalA2_Per|PartSop_IntervalA3_Per|PartSop_IntervalA4_Per|PartSop_IntervalA5_Per|PartSop_IntervalA6_Per|PartSop_IntervalA8_Per")) |>
    names()

  # Features subset
  features_subsets$Sop_features_non_cat <-
    features_subsets$All_sop_features |> select(-all_of(cols_to_drop))
  return(features_subsets)

}

# Compute F1 score function
f1 <- function(y, y_hat, average = c("macro", "weighted")) {

  # Ensure same factor levels
  y <- factor(y)
  y_hat <- factor(y_hat, levels = levels(y))
  classes <- levels(y)
  n <- length(y)
  f1_per_class <- numeric(length(classes))
  weights <- numeric(length(classes))

  # Compute F1 score for each class
  for (i in seq_along(classes)) {

    cls <- classes[i]
    tp <- sum(y_hat == cls & y == cls)
    fp <- sum(y_hat == cls & y != cls)
    fn <- sum(y_hat != cls & y == cls)
    precision <- ifelse(tp + fp == 0, 0, tp / (tp + fp))
    recall <- ifelse(tp + fn == 0, 0, tp / (tp + fn))
    f1 <- ifelse(precision + recall == 0,
                 0,
                 2 * precision * recall / (precision + recall))
    f1_per_class[i] <- f1
    weights[i] <- sum(y == cls) / n

  }

  # Aggregation of F1 scores
  average <- match.arg(average)
  if (average == "macro") {

    return(mean(f1_per_class))

  } else {

    return(sum(weights * f1_per_class))

  }

}

# Bootstrap accuracy
boot_acc <- function(true_labels, pred_labels, n_boot = 1000, seed = 123) {

  set.seed(seed) # It is key to ensure the same bootstrap samples are used
  # for different applications of the function for different models
  n <- length(true_labels)
  acc <- numeric(n_boot)
  for (i in seq_len(n_boot)) {

    indices <- sample(1:n, size = n, replace = TRUE)
    acc[i] <- mean(true_labels[indices] == pred_labels[indices])

  }
  return(acc)

}
```

# Read datasets

```{r, read-data}
# Load csv
df <- read.csv(paste0("Experiment_", experiment, ".csv"),
               stringsAsFactors = TRUE)

# Choose target column based on experiment
target_col <- case_when(
  experiment %in% c(1, 2) ~ "Gender",
  experiment == 3 ~ "Singer_Sex"
)

# Rename target to 'y'
df <- df |>
  rename(y = !!target_col)

# Number of observations per group
df |>
  group_by(y) |>
  summarise(n = n(), .groups = "drop")

# Split into train / test (85% / 15%)
set.seed(123)
split_obj <- initial_split(df, prop = 0.85)
train <- training(split_obj)
test <- testing(split_obj)
```

```{r, recipes-subset}
# Determine subset based ONLY on training data
features_list_train <- get_features_subsets_voice(train)
if (!(subset_name %in% names(features_list_train))) {
  stop("Subset ", subset_name, " not found")
}

# Select features based on subset name
selected_features <- colnames(features_list_train[[subset_name]])
train <- train |>
  select(all_of(selected_features), y)
test <- test |>
  select(all_of(selected_features), y)
dim(train)
dim(test)

# Combined dataset for recipes
combined <- bind_rows(train, test)
```

# Preprocessing recipes

```{r, recipes}
# Preprocessing recipe
preprocess_recipe <- function(data) {
  recipe(y ~ ., data = data) |>
    # Preprocess all predictors BEFORE dummying
    step_zv(all_predictors()) |>
    step_nzv(all_predictors()) |>
    step_filter_missing(all_predictors(), threshold = 0.05) |>
    step_impute_knn(all_predictors(), neighbors = 5) |>
    # Collapse rare factor levels (BEFORE dummy creation)
    step_other(all_nominal_predictors(), threshold = 0.10) |>
    # Preprocess numeric predictors only
    step_corr(all_numeric_predictors(), threshold = 0.80) |>
    step_lincomb(all_numeric_predictors(), max_steps = 20) |>
    step_normalize(all_numeric_predictors()) |>
    # Create dummy variables (AFTER normalization)
    # Use one-hot encoding such that no dummy variable is dropped and we can
    # later drop the "other" level manually (so it becomes the reference level
    # and we avoid hard-to-explain dummies in the final model)
    # step_dummy(all_nominal_predictors(), one_hot = TRUE) |>
    step_dummy(all_nominal_predictors(), one_hot = FALSE) |>
    # Drop the dummies for the "other" level
    step_rm(ends_with("_other")) |>
    # Post-dummy cleanup
    step_zv(all_predictors()) |>
    step_nzv(all_predictors()) |>
    step_lincomb(all_predictors(), max_steps = 20)
}
```

```{r, recipes-bake}
# Train recipe using train data
rec_train <- preprocess_recipe(data = train)
prep_rec_train <- prep(rec_train)

# Bake train data
train_bake <- bake(prep_rec_train, new_data = NULL)
test_train_bake <- bake(prep_rec_train, new_data = test)

# Train recipe using combined data (train + test)
rec_all <- preprocess_recipe(data = combined)
prep_all <- prep(rec_all)

# Fully baked dataset for the final model fit
all_bake <- bake(prep_all, new_data = NULL)

# Export datasets for future reference and use
dir.create("preprocessed_datasets", showWarnings = FALSE)
write.csv(train_bake, paste0("preprocessed_datasets/train_bake_exp", experiment,
                             "_sub_", subset_name, ".csv"),
          row.names = FALSE)
write.csv(test_train_bake, paste0("preprocessed_datasets/test_train_bake_exp",
                                  experiment, "_sub_", subset_name, ".csv"),
          row.names = FALSE)
write.csv(all_bake, paste0("preprocessed_datasets/all_bake_", experiment,
                           "_sub_", subset_name, ".csv"),
          row.names = FALSE)
```

# Lasso logistic

## Fitting

```{r, log1-fit}
# Fit by 10-fold cross-validation
set.seed(42)
train_bake_no_y <- train_bake |>
  select(-y) |>
  as.matrix()
cv1_train <- cv.glmnet(x = train_bake_no_y, y = train_bake$y,
                       family = "binomial", alpha = 1, nfolds = 10)
plot(cv1_train)
```

## Predictions

```{r, log1-pred}
# Predictions
test_train_bake_no_y <- test_train_bake |>
  select(-y) |>
  as.matrix()
y_test_hat_LL <- predict(cv1_train, newx = test_train_bake_no_y,
                         s = "lambda.1se", type = "class")
p_test_hat_LL <- predict(cv1_train, newx = test_train_bake_no_y,
                         s = "lambda.1se", type = "response")
```

## Metrics

```{r, log1-metrics}
# Metrics
acc_LL <- mean(test$y == y_test_hat_LL)
auc_LL <- auc_binary(test$y, p_test_hat_LL)
f1_LL <- f1(test$y, y_test_hat_LL)
data.frame("Accuracy" = acc_LL, "AUC" = auc_LL, "Macro_F1" = f1_LL) |>
  kable(digits = 4)
```

## Relevant coefficients

```{r, log1-coefs}
# Extract nonzero coefficients for lambda.1se and order them by absolute value
coefs1_train <- predict(cv1_train, s = "lambda.1se", type = "coefficients") |>
  as.matrix() |>
  as.data.frame() |>
  rownames_to_column(var = "feature") |>
  rename("coefficient" = "lambda.1se")
coefs1_train |>
  filter(coefficient != 0) |>
  arrange(desc(abs(coefficient))) |>
  kable(digits = 4)
```

# Ridge logistic model

## Fitting

```{r, log0-fit}
# Fit by 10-fold cross-validation
set.seed(42)
cv0_train <- cv.glmnet(x = train_bake_no_y, y = train_bake$y,
                       family = "binomial", alpha = 0, nfolds = 10)
plot(cv0_train)
```

## Predictions

```{r, log0-pred}
# Predictions
y_test_hat_RL <- predict(cv0_train, newx = test_train_bake_no_y,
                         s = "lambda.1se", type = "class")
p_test_hat_RL <- predict(cv0_train, newx = test_train_bake_no_y,
                         s = "lambda.1se", type = "response")
```

## Metrics

```{r, log0-metrics}
# Metrics
acc_RL <- mean(test$y == y_test_hat_RL)
auc_RL <- auc_binary(test$y, p_test_hat_RL)
f1_RL <- f1(test$y, y_test_hat_RL)
data.frame("Accuracy" = acc_RL, "AUC" = auc_RL, "Macro_F1" = f1_RL) |>
  kable(digits = 4)
```

## Relevant coefficients

```{r, log0-coefs}
# Extract coefficients and sort them by absolute value
coefs0_train <- predict(cv0_train, s = "lambda.1se", type = "coefficients") |>
  as.matrix() |>
  as.data.frame() |>
  rownames_to_column(var = "feature") |>
  rename("coefficient" = "lambda.1se")
coefs0_train |>
  filter(coefficient != 0) |>
  arrange(desc(abs(coefficient))) |>
  head(30) |>
  kable(digits = 4)
```

# Random forests

## Fitting

```{r, rf-fit}
#| message: false
# Set seeds for reproducibility in caret with RF
set.seed(42)
n_folds <- 10
n_tune <- 5
rf_seeds <- c(
  lapply(seq_len(n_folds), function(i) rep(i, n_tune)),
  list(999L)
)

# Define 10-fold CV
cv_ctrl <- trainControl(method = "cv", number = 10, classProbs = TRUE,
                        summaryFunction = twoClassSummary,
                        savePredictions = "final", seeds = rf_seeds)

# Train Random Forest with tuning of mtry optimizing AUC
rf_train <- train(x = train_bake_no_y, y = train_bake$y, method = "rf",
                  metric = "ROC", trControl = cv_ctrl, tuneLength = n_tune)
plot(rf_train)
```

## Predictions

```{r, rf-pred}
# Predictions
y_test_hat_RF <- predict(rf_train, newdata = test_train_bake_no_y)
positive_class <- levels(train_bake$y)[2] # "Class 1" is the second factor level
p_test_hat_RF <- predict(rf_train, newdata = test_train_bake_no_y,
                         type = "prob")[, positive_class]
```

## Metrics

```{r, rf-metrics}
# Metrics
acc_RF <- mean(test_train_bake$y == y_test_hat_RF)
auc_RF <- auc_binary(test$y, p_test_hat_RF)
f1_RF <- f1(test$y, y_test_hat_RF)
data.frame("Accuracy" = acc_RF, "AUC" = auc_RF, "Macro_F1" = f1_RF) |>
  kable(digits = 4)
```

## Variables importance

```{r, rf-vars}
# Variables importance
rf_importance <- varImp(rf_train)$importance
rf_importance |>
  rownames_to_column("feature") |>
  arrange(desc(Overall)) |>
  head(30) |>
  kable(digits = 4)
```

# Metrics for all models on test data

```{r, save-predictions}
# Save predictions for all models
dir.create("results/y_hats", showWarnings = FALSE)
datasets_to_save <- list(
  RL = y_test_hat_RL,
  LL = y_test_hat_LL,
  RF = y_test_hat_RF,
  y_true = test$y
)
for (name in names(datasets_to_save)) {

  data_to_write <- datasets_to_save[[name]]

  # If it's a vector, convert to data frame
  if (is.vector(data_to_write)) {
    data_to_write <- data.frame(prediction = data_to_write)
  }

  # Construct filename
  file_name <- paste0("results/y_hats/", name, "_", experiment,
                      "_sub_", subset_name, ".csv")

  # Write CSV
  write.csv(data_to_write, file_name, row.names = FALSE)

}
```

```{r, all-metrics}
# Combine all metrics into a single table
metrics_table <- data.frame(
  Accuracy = c(acc_RL, acc_LL, acc_RF),
  AUC = c(auc_RL, auc_LL, auc_RF),
  Macro_F1 = c(f1_RL, f1_LL, f1_RF),
  Model = c("Ridge Logistic (RL)", "Lasso Logistic (LL)",
            "Random Forest (RF)")
)
metrics_table |>
  kable(digits = 4)
```

# Bootstrap tests for model comparison

```{r, boot}
# Run bootstrapping of accuracy with predictions
B <- 10000
boot_RL <- boot_acc(test$y, y_test_hat_RL, n_boot = B, seed = 123)
boot_LL <- boot_acc(test$y, y_test_hat_LL, n_boot = B, seed = 123)
boot_RF <- boot_acc(test$y, y_test_hat_RF, n_boot = B, seed = 123)
boot <- list(RL = boot_RL, LL = boot_LL, RF = boot_RF)

# Bootstrap accuracy differences between RF and other models
diff_boot_RF_RL <- boot$RF - boot$RL
diff_boot_RF_LL <- boot$RF - boot$LL

# One-sided p-values (H1: RF better than other model)
p_RF_vs_RL <- (1 + sum(diff_boot_RF_RL <= 0)) / (B + 1)
p_RF_vs_LL <- (1 + sum(diff_boot_RF_LL <= 0)) / (B + 1)

# Two-sided 95% confidence intervals for differences
ci_diff_RL <- quantile(diff_boot_RF_RL, probs = c(0.025, 0.975))
ci_diff_LL <- quantile(diff_boot_RF_LL, probs = c(0.025, 0.975))

# Add bootstrap tests to metrics table
metrics_table <- metrics_table |>
  mutate(
    ci_diff_lo = c(ci_diff_RL[1], ci_diff_LL[1], NA),
    ci_diff_up = c(ci_diff_RL[2], ci_diff_LL[2], NA),
    pval_better = c(p_RF_vs_RL, p_RF_vs_LL, NA)
  )
metrics_table |>
  kable(digits = 4)
```

# Comparison with majority class baseline

```{r, mc-baseline}
# Majority Class (MC) prediction baseline - always predict most common class
class_probs <- table(train$y) / length(train$y)
majority_class_train <- names(sort(class_probs, decreasing = TRUE))[1]
y_test_hat_MC <- factor(rep(majority_class_train, length(test$y)),
                        levels = levels(test$y))

# Metrics for majority class baseline
acc_MC <- mean(test$y == y_test_hat_MC)
p_test_hat_MC <- rep(class_probs[majority_class_train], length(test$y))
auc_MC <- auc_binary(test$y, p_test_hat_MC)
f1_MC <- f1(test$y, y_test_hat_MC)

# Bootstrap accuracy for majority class baseline
boot_MC <- boot_acc(test$y, y_test_hat_MC, n_boot = B, seed = 123)

# Bootstrap accuracy differences between majority class baseline and models
diff_boot_RL_MC <- boot_RL - boot_MC
diff_boot_LL_MC <- boot_LL - boot_MC
diff_boot_RF_MC <- boot_RF - boot_MC

# One-sided p-values (H1: model better than majority class baseline)
p_RL_vs_MC <- (1 + sum(diff_boot_RL_MC <= 0)) / (B + 1)
p_LL_vs_MC <- (1 + sum(diff_boot_LL_MC <= 0)) / (B + 1)
p_RF_vs_MC <- (1 + sum(diff_boot_RF_MC <= 0)) / (B + 1)

# Two-sided 95% confidence intervals for differences
ci_diff_RL_MC <- quantile(diff_boot_RL_MC, probs = c(0.025, 0.975))
ci_diff_LL_MC <- quantile(diff_boot_LL_MC, probs = c(0.025, 0.975))
ci_diff_RF_MC <- quantile(diff_boot_RF_MC, probs = c(0.025, 0.975))

# Combine all metrics into a single table
metrics_table_MC <- data.frame(
  Accuracy = c(acc_RL, acc_LL, acc_RF, acc_MC),
  AUC = c(auc_RL, auc_LL, auc_RF, auc_MC),
  Macro_F1 = c(f1_RL, f1_LL, f1_RF, f1_MC),
  ci_diff_lo = c(ci_diff_RL_MC[1], ci_diff_LL_MC[1], ci_diff_RF_MC[1], NA),
  ci_diff_up = c(ci_diff_RL_MC[2], ci_diff_LL_MC[2], ci_diff_RF_MC[2], NA),
  pval_better = c(p_RL_vs_MC, p_LL_vs_MC, p_RF_vs_MC, NA),
  Model = c("Ridge Logistic (RL)", "Lasso Logistic (LL)",
            "Random Forest (RF)", "Majority Class (MC)")
)
metrics_table_MC |>
  kable(digits = 4)

# In a combined table
metrics_table_combined <- metrics_table_MC |>
  mutate(ci_diff_lo_MC = ci_diff_lo, ci_diff_up_MC = ci_diff_up,
         pval_better_MC = pval_better) |>
  mutate(ci_diff_lo_RF = c(metrics_table$ci_diff_lo, NA),
         ci_diff_up_RF = c(metrics_table$ci_diff_up, NA),
         pval_better_RF = c(metrics_table$pval_better, NA)) |>
  select(Model, Accuracy, AUC, Macro_F1,
         ci_diff_lo_RF, ci_diff_up_RF, pval_better_RF,
         ci_diff_lo_MC, ci_diff_up_MC, pval_better_MC)
metrics_table_combined |>
  kable(digits = 4)

# Save table of all metrics + bootstrap tests
dir.create("results/experiments", showWarnings = FALSE)
write.csv(metrics_table_combined, paste0("results/experiments/all_results_",
                                         "Experiment_", experiment,
                                         "_sub_", subset_name, ".csv"),
          row.names = FALSE)
```

\scriptsize

```{r, latex-table}
# Print LaTeX table
latex_table <- metrics_table_combined |>
  mutate(ci_RF = paste0("$(", sprintf("%.4f", ci_diff_lo_RF), ", ",
                        sprintf("%.4f", ci_diff_up_RF), ")$"),
         ci_MC = paste0("$(", sprintf("%.4f", ci_diff_lo_MC), ", ",
                        sprintf("%.4f", ci_diff_up_MC), ")$"),
         p_RF = sprintf("%.4f", pval_better_RF),
         p_MC = sprintf("%.4f", pval_better_MC)) |>
  select(Model, Accuracy, AUC, Macro_F1, ci_RF, p_RF, ci_MC, p_MC)
latex_table |>
  kable(digits = 4)
latex_table |>
  kable(digits = 4, format = "latex", booktabs = TRUE, escape = FALSE) |>
  print()
```

\normalsize

# Ridge logistic for inference and final model on all data (train + test)

```{r, ridge-inf-fit}
# Fit ridge
set.seed(42)
all_bake_no_y <- all_bake |>
  select(-y) |>
  as.matrix()
cv0_all <- cv.glmnet(x = all_bake_no_y, y = all_bake$y, family = "binomial",
                     alpha = 0, nfolds = 10)
plot(cv0_all)
cv0_all
```

```{r, ridge-inf-coefs}
#| out.width: '100%'
#| fig.width: 10
#| fig.height: 10
# Confidence intervals for all coefficients
ci_df <- logistic_ridge_ci(X = all_bake_no_y, y = all_bake$y,
                           beta_hat = coef(cv0_all, s = cv0_all$lambda.1se),
                           lambda = nrow(all_bake) * cv0_all$lambda.1se / 2)

# Filter significant coefficients and order by effect size
ci_df <- ci_df |>
  filter(sign(ci_low * ci_up) > 0) |>
  mutate(
    pct_change = (exp(estimate) - 1) * 100,
    pct_low = (exp(ci_low) - 1) * 100,
    pct_up = (exp(ci_up) - 1) * 100
  ) |>
  mutate(
    term = factor(term, levels = term[order(pct_change, decreasing = FALSE)]
    )
  )
ci_df |>
  select(term, pct_change, pct_low, pct_up, p_value) |>
  arrange(desc(pct_change)) |>
  kable(digits = 2)

# Plot of coefficients in odds
plot_RL <- ggplot(ci_df, aes(x = term, y = pct_change,
                             ymin = pct_low, ymax = pct_up)) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = 2) +
  coord_flip() +
  labs(
    x = NULL,
    y = "Percent change in odds (per 1-unit SD increase in numerical\npredictor or presence of categorical predictor)"
  ) +
  theme_bw()
plot_RL

# Save plot
plot_name <- paste0("plots/plot_rl", "_", experiment,
                    "_sub_", subset_name, ".png")
ggsave(filename = plot_name, plot = plot_RL, width = 8, height = 6, dpi = 300)

# Metrics on all data available, just for reference
y_all_hat_RL <- predict(cv0_all, newx = all_bake_no_y, s = "lambda.1se",
                        type = "class")
p_all_hat_RL <- predict(cv0_all, newx = all_bake_no_y, s = "lambda.1se",
                        type = "response")
acc_all_RL <- mean(all_bake$y == y_all_hat_RL)
auc_all_RL <- auc_binary(all_bake$y, p_all_hat_RL)
f1_all_RL <- f1(all_bake$y, y_all_hat_RL)
glmnet_summary <- data.frame("Accuracy" = acc_all_RL, "AUC" = auc_all_RL,
                             "Macro_F1" = f1_all_RL,
                             "Model" = "Ridge Logistic (RL)")
glmnet_summary |>
  kable(digits = 4)

# Save results
dir.create("results/final_model", showWarnings = FALSE)
write.csv(glmnet_summary, paste0("results/final_model/RL_final_model_results",
                                 "_Experiment_", experiment, "_sub_",
                                 subset_name, ".csv"),
          row.names = FALSE)
```
