---
title: "Gender dynamics experiments"
params:
  experiment: 1
  subset_name: "Sop_features_non_cat"
author: "Eduardo García-Portugués, Carlos Vaquero"
date: '`r Sys.Date()`'
format:
  pdf:
    latex_engine: xelatex
    number_sections: yes
    keep-tex: false
fontsize: 10pt
urlcolor: blue
editor_options:
  chunk_output_type: console
---

```{=html}
<!--

# This report is rendered once per (experiment, subset) combination.
# To reproduce all experiments, run ./experiments.sh from the repository root.
# Alternatively, execute the following block in R:

exps <- 1:3
subsets <- c(
  "Sop_features_non_cat"
)
for (exp in exps) {
  for (s in subsets) {
    message("Rendering for experiment: ", exp, " and subset: ", s)
    quarto::quarto_render(
      input = "Models.qmd",
      execute_params = list(experiment = exp, subset_name = s),
      output_file = paste0("Experiment_", exp,"_", s, ".pdf")
    )
  }
}

--->
```

```{r, setup-params}
#| echo: false
# Report parameters
subset_name <- params$subset_name
experiment <- params$experiment
message("Running experiment ", experiment, " with subset ", subset_name)

# Analysis constants, referenced throughout the report
n_folds <- 10 # Folds for every cross-validation (glmnet and caret)
n_tune <- 5 # Grid size for the random forest mtry tuning
B <- 10000 # Bootstrap replicates for model comparison
level <- 0.95 # Confidence level for all intervals
ci_probs <- c((1 - level) / 2, 1 - (1 - level) / 2)
```

# Load libraries

```{r, libraries}
#| message: false
# Load libraries
library(knitr)
library(dplyr)
library(recipes)
library(rsample)
library(yardstick)
library(glmnet)
library(ggplot2)
library(caret)
library(randomForest)
library(tibble)
sessionInfo()
```

```{r, functions}
#| include: false
#| cache: false
knitr::opts_chunk$set(
  echo = TRUE, cache = FALSE, warning = FALSE, message = FALSE,
  fig.path = paste0("Models_files/Experiment_", params$experiment, "_",
                    params$subset_name, "/")
)

# Compute the AUC of a binary classifier.
auc_binary <- function(y, p_hat) {

  roc_auc_vec(
    truth = y,
    estimate = drop(p_hat),
    event_level = "second"
  )

}

# Confidence intervals for the coefficients of a logistic ridge fit.
#
# X: design matrix (n x p), WITHOUT an intercept column (added internally)
# beta_hat: vector of coefficients (length p + 1), intercept first
# lambda: penalty parameter in "theory" scale: (X'X + 2 * lambda * I)
#         If the input of lambda is that from glmnet, then set
#         lambda = n * lambda_glmnet / 2
# level: confidence level, default 0.95
# m: Bonferroni family size for the slope intervals; NULL (default) gives
#    marginal intervals. The intercept is not a tested hypothesis and always
#    receives the marginal interval.
# Value: a tibble with one row per coefficient, intercept first
#
# Note: the recipe standardizes every baked column (numerics and dummy
# indicators alike) to unit sample sd, so glmnet's internal standardization
# (1 / n variance) rescales all columns by the same sqrt((n - 1) / n) factor
# and the uniform 2 * lambda penalty below matches glmnet's effective
# per-column penalty up to that common, asymptotically negligible factor.
logistic_ridge_ci <- function(X, beta_hat, lambda, level = 0.95, m = NULL) {

  p <- ncol(X)

  # Design matrix with intercept column, matching beta_hat
  X1 <- cbind("(Intercept)" = 1, X)

  # Binomial weights of the Fisher information
  eta_hat <- drop(X1 %*% beta_hat)
  p_hat <- 1 / (1 + exp(-eta_hat))
  w_hat <- p_hat * (1 - p_hat)

  # Asymptotic covariance from Le Cessie & Van Houwelingen (1992, p. 194;
  # https://www.jstor.org/stable/2347628)
  # with the 2 * lambda ridge penalty and the intercept unpenalized.
  # crossprod(X1 * sqrt(w_hat)) avoids the dense n x n weight matrix
  XtWX <- crossprod(X1 * sqrt(w_hat))
  pen <- diag(c(0, rep(2 * lambda, p)))
  XtWX_lambda_inv <- solve(XtWX + pen)
  asymp_var <- XtWX_lambda_inv %*% XtWX %*% XtWX_lambda_inv

  # z-based CIs; slope intervals Bonferroni-adjusted when m is given
  se <- sqrt(diag(asymp_var))
  alpha <- 1 - level
  alpha_slopes <- if (is.null(m)) alpha else alpha / m
  zcrit <- qnorm(1 - c(alpha, rep(alpha_slopes, p)) / 2)
  est <- as.numeric(beta_hat)
  ci_low <- est - zcrit * se
  ci_up <- est + zcrit * se

  # Adjusted p-values are Bonferroni on the slopes, NA for the intercept
  slope_names <- colnames(X) %||% paste0("x", seq_len(p))

  # Return a tibble
  tibble(
    term = c("(Intercept)", slope_names),
    estimate = est,
    std_error = se,
    ci_low = ci_low,
    ci_up = ci_up,
    z_value = estimate / std_error,
    p_value = 2 * pnorm(-abs(z_value)),
    p_value_adj = c(NA, pmin(1, (if (is.null(m)) 1 else m) * p_value[-1]))
  )

}

# Collapse a set of regex alternatives into a single pattern
alt <- function(...) paste(c(...), collapse = "|")

# Windowed, and jSymbolic / music21 columns, plus the non-Soprano parts
drop_general <- c(
  "_win_", "js_.*", "m21.*", "SoundSop", "Count", "Sum", "Notes",
  "SoundingMeasures", "SopI", "SopII", "SopIII", "PartSop_Syllables"
)

# Melodic contour, range, and interval-size descriptors
melody_core <- c(
  "Amb", "Density", "Trimmed", "Semitones", "AbsoluteIntervallic",
  "AscendingIntervals", "AscendingIntervallic", "StepwiseMotionDesc",
  "PartSop_LowestNoteIndex", "PartSop_HighestNoteIndex"
)

# Interval-quality descriptors, added on top of melody_core
melody_quality <- c(
  "PartSop_Degree", "IntervalA", "IntervalM", "IntervalP", "Intervald",
  "Intervalm"
)

# Rhythmic descriptors
rhythm_terms <- c(
  "DottedRhythm", "AverageDuration", "DoubleDottedRhythm", "RhythmInt"
)

# Text-setting descriptors
lyrics_terms <- c("VoicePresence", "SyllabicRatio")

# Categorical columns excluded from the non-categorical subset
categorical_terms <- c(
  "PartSop_Degree", "PartSop_IntervalM", "PartSop_IntervalP",
  "PartSop_Intervald", "PartSop_Intervalm", "PartSop_MeanInterval",
  "PartSop_SmallestIntervalAsc", "PartSop_LargestIntervalDesc",
  "PartSop_IntervalA.2_Per", "PartSop_IntervalA.4_Per",
  "PartSop_IntervalA.5_Per", "PartSop_IntervalA1_Per",
  "PartSop_IntervalA11_Per", "PartSop_IntervalA2_Per",
  "PartSop_IntervalA3_Per", "PartSop_IntervalA4_Per",
  "PartSop_IntervalA5_Per", "PartSop_IntervalA6_Per",
  "PartSop_IntervalA8_Per"
)

# Extract features according to the types of subsets defined.
get_features_subsets_voice <- function(df_total) {

  # 1. Drop unwanted columns by regex
  df_total <- df_total |>
    select(-matches(alt(drop_general)),
           -matches("^PartSop_HighestNote($|_[A-Za-z0-9]+)"),
           -matches("^PartSop_LowestNote($|_[A-Za-z0-9]+)"))

  # 2. Keep only Sop* features (no target)
  df_total <- df_total |>
    select(matches("^Sop|^PartSop")) |>
    # remove only columns like PartSop_HighestNote_G5 and PartSop_LowestNote_C4
    select(-matches("^PartSop_HighestNote_.+|^PartSop_LowestNote_.+"))

  # 3. Feature subset definitions
  features_subsets <- list(
    Melody = df_total |>
      select(matches(alt(melody_core))) |>
      select(-matches("PartSop_RhythmInt")),

    Melody_extended = df_total |>
      select(matches(alt(melody_quality, melody_core))) |>
      select(-matches("PartSop_RhythmInt")),

    Rhythm = df_total |>
      select(matches(alt(rhythm_terms))),

    Lyrics = df_total |>
      select(matches(alt(lyrics_terms))),

    All_sop_features = df_total |>
      select(matches("PartSop"))
  )

  # 4. Non-categorical version
  cols_to_drop <- features_subsets$All_sop_features |>
    select(matches(alt(categorical_terms))) |>
    names()

  # Features subset
  features_subsets$Sop_features_non_cat <-
    features_subsets$All_sop_features |> select(-all_of(cols_to_drop))
  return(features_subsets)

}

# Compute the aggregated F1 score of a classifier
f1 <- function(y, y_hat, average = c("macro", "weighted")) {

  # Ensure same factor levels
  y <- factor(y)
  y_hat <- factor(y_hat, levels = levels(y))
  classes <- levels(y)
  n <- length(y)
  f1_per_class <- numeric(length(classes))
  weights <- numeric(length(classes))

  # Compute F1 score for each class
  for (i in seq_along(classes)) {

    cls <- classes[i]
    tp <- sum(y_hat == cls & y == cls)
    fp <- sum(y_hat == cls & y != cls)
    fn <- sum(y_hat != cls & y == cls)
    precision <- ifelse(tp + fp == 0, 0, tp / (tp + fp))
    recall <- ifelse(tp + fn == 0, 0, tp / (tp + fn))
    f1_cls <- ifelse(precision + recall == 0,
                     0,
                     2 * precision * recall / (precision + recall))
    f1_per_class[i] <- f1_cls
    weights[i] <- sum(y == cls) / n

  }

  # Aggregation of F1 scores
  average <- match.arg(average)
  if (average == "macro") {

    return(mean(f1_per_class))

  } else {

    return(sum(weights * f1_per_class))

  }

}

# Bootstrap distribution of the accuracy of a classifier
boot_acc <- function(true_labels, pred_labels, n_boot = 1000, seed = 123) {

  set.seed(seed) # It is key to ensure the same bootstrap samples are used
  # for different applications of the function for different models
  n <- length(true_labels)
  acc <- numeric(n_boot)
  for (i in seq_len(n_boot)) {

    indices <- sample(1:n, size = n, replace = TRUE)
    acc[i] <- mean(true_labels[indices] == pred_labels[indices])

  }
  return(acc)

}

```

# Read datasets

```{r, read-data}
# Choose target column based on experiment
target_col <- switch(
  as.character(experiment),
  "1" = "Gender",
  "2" = "Gender",
  "3" = "Singer_Sex",
  stop("Experiment ", experiment, " not supported (expected 1, 2 or 3)")
)

# Load csv
data_file <- paste0("Experiment_", experiment, ".csv")
if (!file.exists(data_file)) {

  stop("Data file ", data_file, " not found; run from the repository root")

}
df <- read.csv(data_file, stringsAsFactors = TRUE)

# Rename target to 'y'
df <- df |>
  rename(y = !!target_col)

# Number of observations per group
df |>
  group_by(y) |>
  summarise(n = n(), .groups = "drop")

# Split into train / test (85% / 15%), stratified by y so that both sets
# preserve the class balance of the full dataset
set.seed(123)
split_obj <- initial_split(df, prop = 0.85, strata = y)
train <- training(split_obj)
test <- testing(split_obj)
```

# Corpus vocal ranges

```{r, fig1-vocal-ranges}
#| fig.width: 7.5
#| fig.height: 2.9
# Two-panel boxplots of the arias' vocal ranges: highest and lowest notes by
# premiering singer's sex (arias with a known singer, Experiment_3.csv) and by
# character gender (all arias, Experiment_1.csv). Independent of the current
# experiment, so it is rendered only in the experiment == 1 pass.
if (experiment == 1) {

  read_ranges <- function(file, target, panel, labels) {
    raw <- read.csv(file, stringsAsFactors = FALSE)
    bind_rows(
      tibble(group = labels[raw[[target]]], note = "Highest notes",
             index = raw$PartSop_HighestNoteIndex),
      tibble(group = labels[raw[[target]]], note = "Lowest notes",
             index = raw$PartSop_LowestNoteIndex)
    ) |>
      mutate(panel = panel)
  }
  ranges_df <- bind_rows(
    read_ranges("Experiment_3.csv", "Singer_Sex", "Singer",
                c("Female" = "Female", "Male" = "Male")),
    read_ranges("Experiment_1.csv", "Gender", "Character",
                c("Female" = "Feminine", "Male" = "Masculine"))
  ) |>
    mutate(panel = factor(panel, levels = c("Singer", "Character")))

  # C-note axis: note index 60 = C4, so octave number = index / 12 - 1
  c_breaks <- seq(48, 96, by = 12)
  plot_ranges <- ggplot(ranges_df, aes(x = group, y = index, fill = note)) +
    geom_boxplot(outlier.shape = 1, outlier.size = 0.8) +
    facet_wrap(~ panel, scales = "free_x", strip.position = "bottom") +
    scale_fill_manual(values = c("Highest notes" = "grey35",
                                 "Lowest notes" = "grey85")) +
    scale_y_continuous(breaks = c_breaks,
                       labels = paste0("C", c_breaks / 12 - 1),
                       limits = c(48, 96)) +
    labs(x = NULL, y = "Note", fill = NULL) +
    theme_bw() +
    theme(strip.placement = "outside",
          strip.background = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = c(0.998, 0.995),
          legend.justification = c(1, 1),
          legend.key.size = grid::unit(0.85, "lines"),
          legend.text = element_text(size = 8),
          # Careful with randomForest::margin()
          legend.margin = ggplot2::margin(1, 3, 1, 1),
          legend.background = element_rect(fill = "white", colour = "grey60",
                                           linewidth = 0.3))
  plot_ranges

  # Save plot
  dir.create("plots", showWarnings = FALSE, recursive = TRUE)
  ggsave("plots/singers_vs_characters_ranges.png", plot = plot_ranges,
         width = 7.5, height = 2.9, dpi = 300, bg = "white")

}
```

```{r, recipes-subset}
# Determine subset based ONLY on training data
features_list_train <- get_features_subsets_voice(train)
if (!(subset_name %in% names(features_list_train))) {
  stop("Subset ", subset_name, " not found")
}

# Select features based on subset name
selected_features <- colnames(features_list_train[[subset_name]])
train <- train |>
  select(all_of(selected_features), y)
test <- test |>
  select(all_of(selected_features), y)
dim(train)
dim(test)

# Combined dataset for recipes
combined <- bind_rows(train, test)
```

# Preprocessing recipes

```{r, recipes}
# Preprocessing recipe
preprocess_recipe <- function(data) {
  recipe(y ~ ., data = data) |>
    # Filter and impute all predictors BEFORE dummying.
    step_zv(all_predictors()) |>
    step_nzv(all_predictors()) |>
    step_filter_missing(all_predictors(), threshold = 0.05) |>
    step_impute_knn(all_predictors(), neighbors = 5) |>
    # Collapse rare factor levels (BEFORE dummy creation)
    step_other(all_nominal_predictors(), threshold = 0.10) |>
    # Decorrelate numeric predictors only
    step_corr(all_numeric_predictors(), threshold = 0.80) |>
    step_lincomb(all_numeric_predictors(), max_steps = 20) |>
    # Create dummy variables. With one_hot = TRUE every
    # level surviving step_other() gets an indicator; dropping "_other" below
    # then makes the pooled rare levels the reference
    step_dummy(all_nominal_predictors(), one_hot = TRUE) |>
    # Drop the dummies for the "other" level, making it the reference
    step_rm(ends_with("_other")) |>
    # Post-dummy cleanup
    step_zv(all_predictors()) |>
    step_nzv(all_predictors()) |>
    step_lincomb(all_predictors(), max_steps = 20) |>
    # Standardize LAST so the dummy indicators are standardized too and every
    # column enters glmnet on unit-sd scale
    step_normalize(all_numeric_predictors())
}
```

```{r, recipes-bake}
# Train recipe using train data
rec_train <- preprocess_recipe(data = train)
prep_rec_train <- prep(rec_train)

# Bake train data
train_bake <- bake(prep_rec_train, new_data = NULL)
test_train_bake <- bake(prep_rec_train, new_data = test)

# Train recipe using combined data (train + test)
rec_all <- preprocess_recipe(data = combined)
prep_all <- prep(rec_all)

# Fully baked dataset for the final model fit
all_bake <- bake(prep_all, new_data = NULL)

# Export datasets for future reference and use
dir.create("preprocessed_datasets", showWarnings = FALSE, recursive = TRUE)
write.csv(train_bake, paste0("preprocessed_datasets/train_bake_exp", experiment,
                             "_sub_", subset_name, ".csv"),
          row.names = FALSE)
write.csv(test_train_bake, paste0("preprocessed_datasets/test_train_bake_exp",
                                  experiment, "_sub_", subset_name, ".csv"),
          row.names = FALSE)
write.csv(all_bake, paste0("preprocessed_datasets/all_bake_", experiment,
                           "_sub_", subset_name, ".csv"),
          row.names = FALSE)
```

# Lasso logistic

## Fitting

```{r, log1-fit}
# Fit by 10-fold cross-validation, with folds stratified by y and shared
# between the lasso and ridge fits so that both models see the same partitions
set.seed(42)
train_bake_no_y <- train_bake |>
  select(-y) |>
  as.matrix()
foldid_train <- caret::createFolds(train_bake$y, k = n_folds, list = FALSE)
cv1_train <- cv.glmnet(x = train_bake_no_y, y = train_bake$y,
                       family = "binomial", alpha = 1, foldid = foldid_train)
plot(cv1_train)
```

## Predictions

```{r, log1-pred}
# Predictions
test_train_bake_no_y <- test_train_bake |>
  select(-y) |>
  as.matrix()
y_test_hat_LL <- predict(cv1_train, newx = test_train_bake_no_y,
                         s = "lambda.1se", type = "class")
p_test_hat_LL <- predict(cv1_train, newx = test_train_bake_no_y,
                         s = "lambda.1se", type = "response")
```

## Metrics

```{r, log1-metrics}
# Metrics
acc_LL <- mean(test$y == y_test_hat_LL)
auc_LL <- auc_binary(test$y, p_test_hat_LL)
f1_LL <- f1(test$y, y_test_hat_LL)
data.frame("Accuracy" = acc_LL, "AUC" = auc_LL, "Macro_F1" = f1_LL) |>
  kable(digits = 4)
```

## Relevant coefficients

```{r, log1-coefs}
# Extract nonzero coefficients for lambda.1se and order them by absolute value
coefs1_train <- predict(cv1_train, s = "lambda.1se", type = "coefficients") |>
  as.matrix() |>
  as.data.frame() |>
  rownames_to_column(var = "feature") |>
  rename("coefficient" = "lambda.1se")
coefs1_train |>
  filter(coefficient != 0) |>
  arrange(desc(abs(coefficient))) |>
  kable(digits = 4)
```

# Ridge logistic model

## Fitting

```{r, log0-fit}
# Fit by 10-fold cross-validation, reusing the stratified folds of the lasso
cv0_train <- cv.glmnet(x = train_bake_no_y, y = train_bake$y,
                       family = "binomial", alpha = 0, foldid = foldid_train)
plot(cv0_train)
```

## Predictions

```{r, log0-pred}
# Predictions
y_test_hat_RL <- predict(cv0_train, newx = test_train_bake_no_y,
                         s = "lambda.1se", type = "class")
p_test_hat_RL <- predict(cv0_train, newx = test_train_bake_no_y,
                         s = "lambda.1se", type = "response")
```

## Metrics

```{r, log0-metrics}
# Metrics
acc_RL <- mean(test$y == y_test_hat_RL)
auc_RL <- auc_binary(test$y, p_test_hat_RL)
f1_RL <- f1(test$y, y_test_hat_RL)
data.frame("Accuracy" = acc_RL, "AUC" = auc_RL, "Macro_F1" = f1_RL) |>
  kable(digits = 4)
```

## Relevant coefficients

```{r, log0-coefs}
# Extract coefficients and sort them by absolute value
coefs0_train <- predict(cv0_train, s = "lambda.1se", type = "coefficients") |>
  as.matrix() |>
  as.data.frame() |>
  rownames_to_column(var = "feature") |>
  rename("coefficient" = "lambda.1se")
coefs0_train |>
  filter(coefficient != 0) |>
  arrange(desc(abs(coefficient))) |>
  head(30) |>
  kable(digits = 4)
```

# Random forests

## Fitting

```{r, rf-fit}
#| message: false
# Set seeds for reproducibility in caret with RF. caret expects a list of
# length n_folds + 1: one integer vector per fold (as long as the tuning grid)
# plus a final seed for the fit on the full training set.
set.seed(42)
rf_seeds <- c(
  lapply(seq_len(n_folds), function(i) rep(i, n_tune)),
  list(999L)
)

# Define the cross-validation scheme
cv_ctrl <- trainControl(method = "cv", number = n_folds, classProbs = TRUE,
                        summaryFunction = twoClassSummary,
                        savePredictions = "final", seeds = rf_seeds)

# Train Random Forest with tuning of mtry optimizing AUC
rf_train <- train(x = train_bake_no_y, y = train_bake$y, method = "rf",
                  metric = "ROC", trControl = cv_ctrl, tuneLength = n_tune)
plot(rf_train)
```

## Predictions

```{r, rf-pred}
# Predictions
y_test_hat_RF <- predict(rf_train, newdata = test_train_bake_no_y)
positive_class <- levels(train_bake$y)[2] # "Class 1" is the second factor level
p_test_hat_RF <- predict(rf_train, newdata = test_train_bake_no_y,
                         type = "prob")[, positive_class]
```

## Metrics

```{r, rf-metrics}
# Metrics
acc_RF <- mean(test$y == y_test_hat_RF)
auc_RF <- auc_binary(test$y, p_test_hat_RF)
f1_RF <- f1(test$y, y_test_hat_RF)
data.frame("Accuracy" = acc_RF, "AUC" = auc_RF, "Macro_F1" = f1_RF) |>
  kable(digits = 4)
```

## Variables importance

```{r, rf-vars}
# Variables importance
rf_importance <- varImp(rf_train)$importance
rf_importance |>
  rownames_to_column("feature") |>
  arrange(desc(Overall)) |>
  head(30) |>
  kable(digits = 4)
```

# Metrics for all models on test data

```{r, save-predictions}
# Save predictions for all models
dir.create("results/y_hats", showWarnings = FALSE, recursive = TRUE)
datasets_to_save <- list(
  RL = y_test_hat_RL,
  LL = y_test_hat_LL,
  RF = y_test_hat_RF,
  y_true = test$y
)
for (name in names(datasets_to_save)) {

  # Coerce to a one-column data frame with a stable column name. Note that
  # is.vector() is FALSE for both the glmnet class-prediction matrices and the
  # y_true factor, so the labels have to be extracted with as.vector().
  data_to_write <- data.frame(
    prediction = as.vector(datasets_to_save[[name]])
  )

  # Construct filename
  file_name <- paste0("results/y_hats/", name, "_", experiment,
                      "_sub_", subset_name, ".csv")

  # Write CSV
  write.csv(data_to_write, file_name, row.names = FALSE)

}
```

```{r, all-metrics}
# Combine all metrics into a single table
metrics_table <- data.frame(
  Accuracy = c(acc_RL, acc_LL, acc_RF),
  AUC = c(auc_RL, auc_LL, auc_RF),
  Macro_F1 = c(f1_RL, f1_LL, f1_RF),
  Model = c("Ridge Logistic (RL)", "Lasso Logistic (LL)",
            "Random Forest (RF)")
)
metrics_table |>
  kable(digits = 4)
```

# Bootstrap tests for model comparison

```{r, boot}
# Run bootstrapping of accuracy with predictions
boot_RL <- boot_acc(test$y, y_test_hat_RL, n_boot = B, seed = 123)
boot_LL <- boot_acc(test$y, y_test_hat_LL, n_boot = B, seed = 123)
boot_RF <- boot_acc(test$y, y_test_hat_RF, n_boot = B, seed = 123)
boot <- list(RL = boot_RL, LL = boot_LL, RF = boot_RF)

# Bootstrap accuracy differences between RF and other models
diff_boot_RF_RL <- boot$RF - boot$RL
diff_boot_RF_LL <- boot$RF - boot$LL

# One-sided p-values (H1: RF better than other model)
p_RF_vs_RL <- (1 + sum(diff_boot_RF_RL <= 0)) / (B + 1)
p_RF_vs_LL <- (1 + sum(diff_boot_RF_LL <= 0)) / (B + 1)

# Two-sided confidence intervals for differences
ci_diff_RL <- quantile(diff_boot_RF_RL, probs = ci_probs)
ci_diff_LL <- quantile(diff_boot_RF_LL, probs = ci_probs)

# Add bootstrap tests to metrics table
metrics_table <- metrics_table |>
  mutate(
    ci_diff_lo = c(ci_diff_RL[1], ci_diff_LL[1], NA),
    ci_diff_up = c(ci_diff_RL[2], ci_diff_LL[2], NA),
    pval_better = c(p_RF_vs_RL, p_RF_vs_LL, NA)
  )
metrics_table |>
  kable(digits = 4)
```

# Comparison with majority class baseline

```{r, mc-baseline}
# Majority Class (MC) prediction baseline - always predict most common class
class_probs <- table(train$y) / length(train$y)
majority_class_train <- names(sort(class_probs, decreasing = TRUE))[1]
y_test_hat_MC <- factor(rep(majority_class_train, length(test$y)),
                        levels = levels(test$y))

# Metrics for majority class baseline
acc_MC <- mean(test$y == y_test_hat_MC)
p_test_hat_MC <- rep(class_probs[majority_class_train], length(test$y))
auc_MC <- auc_binary(test$y, p_test_hat_MC)
f1_MC <- f1(test$y, y_test_hat_MC)

# Bootstrap accuracy for majority class baseline
boot_MC <- boot_acc(test$y, y_test_hat_MC, n_boot = B, seed = 123)

# Bootstrap accuracy differences between majority class baseline and models
diff_boot_RL_MC <- boot_RL - boot_MC
diff_boot_LL_MC <- boot_LL - boot_MC
diff_boot_RF_MC <- boot_RF - boot_MC

# One-sided p-values (H1: model better than majority class baseline)
p_RL_vs_MC <- (1 + sum(diff_boot_RL_MC <= 0)) / (B + 1)
p_LL_vs_MC <- (1 + sum(diff_boot_LL_MC <= 0)) / (B + 1)
p_RF_vs_MC <- (1 + sum(diff_boot_RF_MC <= 0)) / (B + 1)

# Two-sided confidence intervals for differences
ci_diff_RL_MC <- quantile(diff_boot_RL_MC, probs = ci_probs)
ci_diff_LL_MC <- quantile(diff_boot_LL_MC, probs = ci_probs)
ci_diff_RF_MC <- quantile(diff_boot_RF_MC, probs = ci_probs)

# Combine all metrics into a single table
metrics_table_MC <- data.frame(
  Accuracy = c(acc_RL, acc_LL, acc_RF, acc_MC),
  AUC = c(auc_RL, auc_LL, auc_RF, auc_MC),
  Macro_F1 = c(f1_RL, f1_LL, f1_RF, f1_MC),
  ci_diff_lo = c(ci_diff_RL_MC[1], ci_diff_LL_MC[1], ci_diff_RF_MC[1], NA),
  ci_diff_up = c(ci_diff_RL_MC[2], ci_diff_LL_MC[2], ci_diff_RF_MC[2], NA),
  pval_better = c(p_RL_vs_MC, p_LL_vs_MC, p_RF_vs_MC, NA),
  Model = c("Ridge Logistic (RL)", "Lasso Logistic (LL)",
            "Random Forest (RF)", "Majority Class (MC)")
)
metrics_table_MC |>
  kable(digits = 4)

# In a combined table
metrics_table_combined <- metrics_table_MC |>
  mutate(ci_diff_lo_MC = ci_diff_lo, ci_diff_up_MC = ci_diff_up,
         pval_better_MC = pval_better) |>
  mutate(ci_diff_lo_RF = c(metrics_table$ci_diff_lo, NA),
         ci_diff_up_RF = c(metrics_table$ci_diff_up, NA),
         pval_better_RF = c(metrics_table$pval_better, NA)) |>
  select(Model, Accuracy, AUC, Macro_F1,
         ci_diff_lo_RF, ci_diff_up_RF, pval_better_RF,
         ci_diff_lo_MC, ci_diff_up_MC, pval_better_MC)
metrics_table_combined |>
  kable(digits = 4)

# Save table of all metrics + bootstrap tests
dir.create("results/experiments", showWarnings = FALSE, recursive = TRUE)
write.csv(metrics_table_combined, paste0("results/experiments/all_results_",
                                         "Experiment_", experiment,
                                         "_sub_", subset_name, ".csv"),
          row.names = FALSE)
```

\scriptsize

```{r, latex-table}
# Print LaTeX table
latex_table <- metrics_table_combined |>
  mutate(ci_RF = paste0("$(", sprintf("%.4f", ci_diff_lo_RF), ", ",
                        sprintf("%.4f", ci_diff_up_RF), ")$"),
         ci_MC = paste0("$(", sprintf("%.4f", ci_diff_lo_MC), ", ",
                        sprintf("%.4f", ci_diff_up_MC), ")$"),
         p_RF = sprintf("%.4f", pval_better_RF),
         p_MC = sprintf("%.4f", pval_better_MC)) |>
  select(Model, Accuracy, AUC, Macro_F1, ci_RF, p_RF, ci_MC, p_MC)
latex_table |>
  kable(digits = 4)
latex_table |>
  kable(digits = 4, format = "latex", booktabs = TRUE, escape = FALSE) |>
  print()
```

\normalsize

# Ridge logistic for inference and final model on all data (train + test)

```{r, ridge-inf-fit}
# Fit ridge
set.seed(42)
all_bake_no_y <- all_bake |>
  select(-y) |>
  as.matrix()
foldid_all <- caret::createFolds(all_bake$y, k = n_folds, list = FALSE)
cv0_all <- cv.glmnet(x = all_bake_no_y, y = all_bake$y, family = "binomial",
                     alpha = 0, foldid = foldid_all)
plot(cv0_all)
cv0_all
```

```{r, ridge-inf-coefs}
#| out.width: '100%'
#| fig.width: 10
#| fig.height: 10
# Bonferroni-corrected confidence intervals for all coefficients; the family
# is the p slopes of this case study (the intercept keeps a marginal CI)
p_slopes <- ncol(all_bake_no_y)
ci_df_all <- logistic_ridge_ci(X = all_bake_no_y,
                               beta_hat = coef(cv0_all, s = cv0_all$lambda.1se),
                               lambda = nrow(all_bake) * cv0_all$lambda.1se / 2,
                               level = level, m = p_slopes)

# Effect sizes as percent changes in odds, exported to CSV for the paper
ci_df_all <- ci_df_all |>
  mutate(
    pct_change = (exp(estimate) - 1) * 100,
    pct_low = (exp(ci_low) - 1) * 100,
    pct_up = (exp(ci_up) - 1) * 100
  )
dir.create("results/final_model", showWarnings = FALSE, recursive = TRUE)
write.csv(ci_df_all, paste0("results/final_model/ci_df_Experiment_",
                            experiment, "_sub_", subset_name, ".csv"),
          row.names = FALSE)

# Significant slopes with and without the Bonferroni correction, for reference
slopes_df <- ci_df_all |>
  filter(term != "(Intercept)")
zcrit_marg <- qnorm(1 - (1 - level) / 2)
data.frame(
  Rule = c("Unadjusted (marginal CI excludes 0)",
           "Bonferroni (simultaneous CI excludes 0)"),
  N_selected = c(sum(abs(slopes_df$z_value) > zcrit_marg),
                 sum(sign(slopes_df$ci_low * slopes_df$ci_up) > 0)),
  N_total = nrow(slopes_df)
) |>
  kable()

# Filter significant slopes (corrected CI excluding zero), order by effect size
ci_df <- slopes_df |>
  filter(sign(ci_low * ci_up) > 0) |>
  mutate(
    term = factor(term, levels = term[order(pct_change, decreasing = FALSE)]
    )
  )
ci_df |>
  select(term, pct_change, pct_low, pct_up, p_value, p_value_adj) |>
  arrange(desc(pct_change)) |>
  kable(digits = 4)

# Plot of coefficients in odds
plot_RL <- ggplot(ci_df, aes(x = term, y = pct_change,
                             ymin = pct_low, ymax = pct_up)) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = 2) +
  coord_flip() +
  labs(
    x = NULL,
    y = paste0("Percent change in odds (per 1-SD increase in each predictor;",
               "\ninterval-category indicators standardized as well)")
  ) +
  theme_bw()
plot_RL

# Save plot
dir.create("plots", showWarnings = FALSE, recursive = TRUE)
plot_name <- paste0("plots/plot_rl", "_", experiment,
                    "_sub_", subset_name, ".png")
ggsave(filename = plot_name, plot = plot_RL, width = 7.5, height = 6,
       dpi = 300)

# Metrics on all data available (in-sample), just for reference
y_all_hat_RL <- predict(cv0_all, newx = all_bake_no_y, s = "lambda.1se",
                        type = "class")
p_all_hat_RL <- predict(cv0_all, newx = all_bake_no_y, s = "lambda.1se",
                        type = "response")
acc_all_RL <- mean(all_bake$y == y_all_hat_RL)
auc_all_RL <- auc_binary(all_bake$y, p_all_hat_RL)
f1_all_RL <- f1(all_bake$y, y_all_hat_RL)
glmnet_summary <- data.frame("Accuracy" = acc_all_RL, "AUC" = auc_all_RL,
                             "Macro_F1" = f1_all_RL,
                             "Model" = "Ridge Logistic (RL)",
                             "Evaluation" = "In-sample (train + test)")
glmnet_summary |>
  kable(digits = 4)

# Save results
dir.create("results/final_model", showWarnings = FALSE, recursive = TRUE)
write.csv(glmnet_summary, paste0("results/final_model/RL_final_model_results",
                                 "_Experiment_", experiment, "_sub_",
                                 subset_name, ".csv"),
          row.names = FALSE)
```
